---
title: "Day 1: 100 Days of Coding"
output: html_notebook
---

I asked ChatGPT to generate a 100 days of code challenge, alternating between using R and Python. 
-----------------------------------------------------------------------------
-----------------------------------------------------------------------------
-----------------------------------------------------------------------------
Day 1:
-----------------------------------------------------------------------------
Project: Write an R script to generate random data and create a basic plot


I will start with creating a dataframe of random data. I will use sample() to randomly sample from 1 to 1000 and will let a value be sampled more than once. I want to create a dataframe of 20 observations across two variables. 
```{r}
data <- data.frame(Var1 = sample(1:1000, size = 20, replace = TRUE), 
                   Var2 = sample(1:1000, size = 20, replace = TRUE))
```
I wanted the data to have a large range with 20 observations in each variable. 

I will now examine the data:
```{r}
head(data)
summary(data)
```
Now, to do a simple scatter plot, using Base-R, to explore the data:
```{r}
plot(data)
```
Now i want to explore the distribution of the data using a historgram:
```{r}
library(ggplot2)
ggplot(data, aes(Var1)) +
  geom_histogram()

ggplot(data, aes(Var2)) +
  geom_histogram()
```

Something that would be interesting to try would be to standardize the data:
```{r}
data$Var1 <- scale(data$Var1)
data$Var2 <- scale(data$Var2)
summary(data)
```
I will now look at various plots for this data, starting with a scatter plot:

```{r}
ggplot(data, aes(Var1, Var2)) +
  geom_point()
```
Density plot for Var1
```{r}
ggplot(data, aes(Var1))+
  geom_density()
```
Density plot for Var2
```{r}
ggplot(data, aes(Var2)) +
  geom_density()
```
This was a rather simple and easy exercise that incorporates: 1)generating random numbers for data and 2) plotting the data

This was a great warm-up to continue with the 100 days of coding. 

