---
title: "Day 3: Descriptive Stats"
output: html_notebook
---

Project:Write an R script to perform descriptive statistics on a dataset

For this project, I will use the iris dataset, as it is easily accessable. 

```{r}
data <- iris
```
I will first want to look at a brief summary of the data set:
```{r}
summary(iris)
```
For further exploration, I will open the dplyr and ggplot2 libraries for more functions
```{r}
library(dplyr)
library(ggplot2)
```
Next I want to explore the distribution of the variables
```{r}
ggplot(data, aes(Sepal.Length)) +
  geom_histogram() +
  labs(title = "Sepal Length")
```
```{r}
ggplot(data, aes(Sepal.Width)) +
  geom_histogram() +
  labs(title = "Sepal Width")
```
```{r}
ggplot(data, aes(Petal.Width)) +
  geom_histogram() +
  labs(title = "Petal Width")
```
```{r}
ggplot(data, aes(Petal.Length)) +
  geom_histogram() +
  labs(title = "Petal Length")
```
Now I will explore the connection between variables, starting with petal length and sepal length. I will start with a correlation table:
```{r}
cor(data[,1:4])
```
According to this, there is a strong correlation between Petal length and width and sepal length. I wan to visualize this now, starting with both length variables: 
```{r}
ggplot(data, aes(Sepal.Length, Petal.Length, color = Species)) +
  geom_point()
```
And now, looking at Petal Width vs sepal length
```{r}
ggplot(data, aes(Sepal.Length, Petal.Width, color = Species)) +
  geom_point()
```
Looking at these graphs, it appears that there is a connection between petal and sepal length but appears to be more so present for versicolor and virginica and not so much for setosa and is more so present when comparing sepal length to petal length. 

```{r}
ggplot(data, aes(Sepal.Length, Petal.Length, color = Species)) +
  geom_point() +
  geom_smooth(method = "lm")
```
Adding a trend line, demonstrates this further with petal and sepal length more connected with versicolor and virginica. 




